### 2020-06-26 manuscript/manuscript.rmd
linted code, did not rerun kernel

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300,
	cache=FALSE
	)
```

@@

```{r preliminaries}
	library( tidyverse )
	library( magrittr )
	library( viridis )
	library( knitr )
	library( gridExtra )
	library( ggpubr )
	library( grid )
	library( gtable )
	theme_set(theme_pubr())
	source( "functions.R" )

	# A colorblind friendly palette from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette:
	cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

@@

taxon_rectangles$matrix[taxon_rectangles$matrix == "Simion2017_subsampled"] <- "Simion2017"
taxon_rectangles$matrix[taxon_rectangles$matrix == "Hejnol2009_subsampled"] <- "Hejnol2009"
#clade_colors = c("#e31a1c","#fb9a99","#fb9a99","#1e90ff","#E69F00","#009E73","#1f78b4","#cab2d6","#c71585")

clade_colors = c('#e6194b', '#f032e6', '#ffe119', '#4363d8', "#E69F00","#009E73","#82d1f4", '#e6beff', '#911eb4')

taxon_rectangles= taxon_rectangles %>% filter(matrix!="Nosenko2013_nonribo_9187") %>% filter(matrix!="Whelan2017_strict") %>% filter(matrix!="Whelan2015_D1") %>% filter(matrix!="Whelan2015_D10") %>% filter(matrix!="Nosenko2013_ribo_11057")%>% filter(matrix!="Borowiec2015_Best108")

taxon_rectangles$matrix= factor(taxon_rectangles$matrix, levels=c("Dunn2008", "Hejnol2009", "Philippe2009", "Nosenko2013_ribo_14615", "Ryan2013_est", "Moroz2014_3d", "Whelan2015_D20", "Borowiec2015_Total1080" , "Chang2015", "Whelan2017_full", "Simion2017"))

ggplot(data=taxon_rectangles) +
	scale_x_continuous(name="Gene sampling") +
	scale_y_continuous(name="Taxa sampling") +
	geom_rect( mapping=aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=clade )) +
	#scale_fill_manual(guide = guide_legend(reverse=TRUE))+
 scale_fill_manual(values = clade_colors, guide = guide_legend(reverse=TRUE))+
	#scale_fill_discrete(guide = guide_legend(reverse=TRUE, legend = "bottom right")) +
	facet_wrap( ~ matrix ) + theme_classic()

@@

analyses_published_processed = 
  analyses_published %>%
  dplyr::filter( !grepl( 'Recoding', model_combined ) ) %>%
  dplyr::rename( model = model_combined ) %>%
  select( model, clade, final, inference, manuscript, matrix ) %>%
  dplyr::rename( result = final ) %>%
  mutate( status = "published" )

@@

analyses_processed = 
  rbind( analyses_published_processed, analyses_new_processed ) %>%
  # mutate(software=recode( software, `PhyloBayes MPI`="PhyloBayes") ) %>%
  mutate(model=recode(model,
                    `GTR20`="GTR",
                    `CAT+F81`="Poisson+CAT",
                    `Poisson + CAT`="Poisson+CAT",
                    `GTR + CAT`="GTR+CAT")) %>%
  mutate( 
    model = factor( model, 
      levels=c("WAG", "LG", "GTR", "data partitioning", "Poisson+C60", "WAG+C60", "LG+C60", "Poisson+CAT", "GTR+CAT") )) %>%
  mutate( status = factor(status, levels = c("published", "new")))

p = analyses_processed %>%
	ggplot( aes(x=model, y=clade, col=result, shape=inference) ) +
    facet_wrap( ~ status ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + 
    theme(axis.text.x = element_text(angle=45, hjust=1))+
    xlab("Model") + 
    ylab("Clade")

@@

# Simion2017 is the only Holozoa analysis to support Porifera-sister
analyses_processed %>% filter(status=="published") %>% filter(clade=="Holozoa") %>% filter(result=="Porifera-sister")

# Check out the CAT analyses that supported Ctenophora sister
analyses_processed %>% filter(status=="published") %>% filter(result=="Ctenophora-sister") %>% filter( grepl("CAT", model))

@@

# See results (and other fields) for new Cxx analyses, all are Ctenophora-sister with exception of Moroz2014_3d and Nosenko matrices
analyses_processed %>% filter(status=="new") %>% filter( grepl("C\\d\\d",model) ) %>% print(n=40)

@@

Philippe2009=analyses_sensitive %>% filter(matrix == "Philippe2009_only_choanozoa" )
Philippe2009$model = factor(Philippe2009$model, levels=c("Poisson+CAT60" , "Poisson+CAT90","Poisson+CAT120","Poisson+CAT150","Poisson+CAT180"))
p1<-ggplot(Philippe2009, aes(model))+
  geom_line(aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group=1), color=("#E69F00")) + geom_point(aes(y = `support_ctenophora_sister` ), color=("#E69F00"))+
  geom_line(aes(y = `support_porifera_sister`, colour = "Porifera-sister",group=2), color=("#009E73")) + geom_point(aes(y = `support_porifera_sister` ), color=("#009E73"))+
  scale_fill_manual(values=c("#E69F00", "#ebd196")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(x ="Number of CAT substitutional categories", y = "Support") + ggtitle("Philippe2009_Choanimalia")

Philippe2009=analyses_sensitive %>% filter(matrix == "Philippe2009_only_Holozoa")
Philippe2009$model = factor(Philippe2009$model, levels=c("Poisson+CAT270", "Poisson+CAT360", "Poisson+CAT400","Poisson+CAT440","Poisson+CAT480"))
p2<-ggplot(Philippe2009, aes(model))+
  geom_line(aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group=1), color=("#E69F00")) + geom_point(aes(y = `support_ctenophora_sister` ), color=("#E69F00"))+
  geom_line(aes(y = `support_porifera_sister`, colour = "Porifera-sister",group=2), color=("#009E73")) + geom_point(aes(y = `support_porifera_sister` ), color=("#009E73"))+
  scale_fill_manual(values=c("#E69F00", "#ebd196")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(x ="Number of CAT substitutional categories", y = "Support") + ggtitle("Philippe2009_Holozoa")

Whelan2017_full_only_choanozoa = analyses_sensitive %>% filter(matrix == "Whelan2017_full_only_choanozoa")
Whelan2017_full_only_choanozoa$model = factor(Whelan2017_full_only_choanozoa$model, levels=c("Poisson+CAT60" ,"Poisson+CAT340","Poisson+CAT380","Poisson+CAT420","Poisson+CAT460"))
p3<-ggplot(Whelan2017_full_only_choanozoa, aes(model))+
  geom_line(aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group=1), color=("#E69F00")) + geom_point(aes(y = `support_ctenophora_sister` ), color=("#E69F00"))+
  geom_line(aes(y = `support_porifera_sister`, colour = "Porifera-sister",group=1), color=("#009E73")) + geom_point(aes(y = `support_porifera_sister` ), color=("#009E73"))+
  scale_fill_manual(values=c("#E69F00", "#ebd196")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(x ="Number of CAT substitutional categories", y = "Support") + ggtitle("Whelan2017_full_Choanozoa")

Whelan2017_strict = analyses_sensitive %>% filter(matrix == "Whelan2017_strict" )
Whelan2017_strict$model = factor(Whelan2017_strict$model, levels=c("Poisson+CAT60" , "Poisson+CAT70","Poisson+CAT80","Poisson+CAT90","Poisson+CAT100","Poisson+CAT110", "Poisson+CAT120","Poisson+CAT150","Poisson+CAT180", "Poisson+CAT360"))
p4<-ggplot(Whelan2017_strict, aes(model))+
  geom_line(aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group=1), color=("#E69F00")) + geom_point(aes(y = `support_ctenophora_sister` ), color=("#E69F00"))+
  geom_line(aes(y = `support_porifera_sister`, colour = "Porifera-sister",group=1), color=("#009E73")) + geom_point(aes(y = `support_porifera_sister` ), color=("#009E73"))+
  scale_fill_manual(values=c("#E69F00", "#ebd196")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(x ="Number of CAT substitutional categories", y = "Support")+ ggtitle("Whelan2017_strict")

ggarrange(ggarrange(p1, p2, ncol = 2, labels = c("A", "B")),nrow = 3,
          p3,p4, labels = c("C", "D"))
```
**Fig. 4.**  Sensitive analyses with representative data matrices analyzed by different number of substitutionary categories (nCAT) in PhyloBayes. Statistical support values (posterior probabilities) were obtained from three data matrcies using the site-heterogeneous Poisson-CAT model with different categories” (A) Phlippe2009_Choanozoa; (B) Philippe2009_Holozoa; (C) Whelan2017_full_Choanozoa; (D) Whelan2017_strict. Statistical support for Ctenophora-sister and Porifera-sister is indicated in orange and green, respectively. Support values are from the sensitive analyses are shown in Supplementary table XX.

@@

**Extended Data Table 2.** The models selected by ModelFinder for each matrix.
```{r table_modelfinder}
analyses_new %>%
	filter(modelfinder==TRUE) %>%
	select( matrix, clade, result, model_summary ) %>%
	kable()

@@

matrix_overlap_rectangles =
	lapply(sequence_matrices[no_subsets], function(x) lapply(sequence_matrices[no_subsets], function(y) overlap_rects(x,y))) %>%
	unlist(recursive=FALSE) %>%
	bind_rows()

ggplot(data=matrix_overlap_rectangles) +
	scale_x_continuous(name="Genes") +
	scale_y_continuous(name="Species") +
	geom_rect( mapping=aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=MSA ), alpha=0.5) +
	facet_grid( matrix_1 ~ matrix_2, scales="free" ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text.y = element_text(angle = 0), strip.text.x = element_text(angle = 60))

@@

ribo$matrix = factor( ribo$matrix, levels=c("Dunn2008", "Philippe2009", "Hejnol2009", "Nosenko2013_nonribo_9187","Nosenko2013_ribo_11057","Nosenko2013_ribo_14615","Ryan2013_est", "Moroz2014_3d", "Chang2015", "Borowiec2015_Best108" , "Borowiec2015_Total1080", "Whelan2015_D1", "Whelan2015_D10", "Whelan2015_D20", "Simion2017", "Whelan2017_full", "Whelan2017_strict") )

p1 = ggplot(ribo, aes(x=matrix, y=number_of_genes, fill=gene_category) ) +geom_bar(stat = "identity")  + 	theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1)) + labs(x ="Matrix", y = "Number of genes")  + scale_fill_manual(values=c("grey", "black"), labels=c("Non-ribosomal genes", "Ribosomal genes"))

busco$matrix = factor( busco $matrix, levels=c("Dunn2008", "Philippe2009", "Hejnol2009", "Nosenko2013_nonribo_9187","Nosenko2013_ribo_11057","Nosenko2013_ribo_14615","Ryan2013_est", "Moroz2014_3d", "Chang2015", "Borowiec2015_Best108" , "Borowiec2015_Total1080", "Whelan2015_D1", "Whelan2015_D10", "Whelan2015_D20", "Simion2017", "Whelan2017_full", "Whelan2017_strict") )
busco$gene_category = factor(busco$gene_category, levels=c("non_busco", "busco"))
p2 = ggplot(busco, aes(x=matrix, y=number_of_genes, fill=gene_category) ) +geom_bar(stat = "identity")  + 	theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1)) + labs(x ="Matrix", y = "Number of genes")  + scale_fill_manual(values=c("grey", "black"), labels=c("Non-BUSCO genes", "BUSCO genes"))

@@

```{r figure_phylogenetic_signal, fig.width=6, fig.height=12}
# fig5_bars <- au_tests %>%
au_tests %>%
  unite(Matrix, c(manuscript, matrix, model)) %>%
  group_by(Matrix, result) %>%
  count() %>%
  group_by(Matrix) %>%
  mutate(percent = prop.table(n)) %>%

ggplot(aes(x = Matrix, y = percent, fill=result)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values=c("#E69F00", "#ebd196","#a2e8d5","#009E73")) +
  coord_flip() +
  theme(axis.text.y=element_blank(), axis.title.y=element_blank())

#fig5_tbl <- au_tests %>%
#select(c(manuscript, matrix, model)) %>%
#distinct()

#fig_5_ylabs_grob <- tableGrob(fig5_tbl)
#fig_5_ylabs$heights = unit(rep(1/nrow(fig_5_ylabs), nrow(fig_5_ylabs)), "npc")
#fig_5_ylabs$widths = unit.pmax(fig_5_ylabs$widths, unit(2, "lines"))

#fig5_bars_grob <- ggplotGrob(fig5_bars)
#fig5 <- gtable_add_cols(fig5, sum(fig_5_ylabs$widths), pos=0)
#fig5 <- gtable_add_grob(fig5, grobs = fig_5_ylabs, t=6, l=ncol(fig5), b=6, r=ncol(fig5))
#fig5 <- gtable_add_grob(fig5, grobs = fig_5_ylabs, t=6, l=1, b=6, r=ncol(fig5))

#grob_grid <- matrix(list(fig_5_ylabs_grob, fig5_bars_grob), nrow = 1)
# https://cran.r-project.org/web/packages/gridExtra/vignettes/gtable.html

# fig5 <- gtable_matrix(name = "demo", grobs = grob_grid,
#                    widths = unit(c(4, 4), "cm"),
#                    heights = unit(c(2, 5), c("in", "lines")))
#
# grid.newpage()
# grid.draw(fig5)
```
**Extended Data Fig. 3** The distribution of phylogenetic signal for two alternative topological hypotheses on the earliest-branching animal lineage with different models and outgroup choice in Philippe2009, Ryan2013 and Whelan2017_full matrices (with two outgroup sampling: Choanozoa and full matrices). The two alternative topological hypotheses are: Ctenophora-sister; T1 (Orange);Porifera-sister T2 (Green). By calculating gene-wise log-likelihood scores between T1 and T2 for every gene, we considered a gene with a log-likelihood difference of two as a gene with strong (lnl > 2) or weak (lnl < 2) phylogenetic signal for each hypothesis (Smith et al. 2020).

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dpi = 300,
  cache = FALSE
  )
```

@@

```{r preliminaries}
  library( tidyverse )
  library( magrittr )
  library( viridis )
  library( knitr )
  library( gridExtra )
  library( ggpubr )
  library( grid )
  library( gtable )

  theme_set(theme_pubr())
  source( "functions.R" )

  # A colorblind friendly palette from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette:
  cbPalette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

@@

taxon_rectangles$matrix[taxon_rectangles$matrix == "Simion2017_subsampled"] = "Simion2017"
taxon_rectangles$matrix[taxon_rectangles$matrix == "Hejnol2009_subsampled"] = "Hejnol2009"

clade_colors = c("#e6194b", "#f032e6", "#ffe119", "#4363d8", "#E69F00", "#009E73", "#82d1f4", "#e6beff", "#911eb4")

taxon_rectangles = 
  taxon_rectangles %>% 
  filter(matrix != "Nosenko2013_nonribo_9187") %>% 
  filter(matrix != "Whelan2017_strict") %>% 
  filter(matrix != "Whelan2015_D1") %>% 
  filter(matrix != "Whelan2015_D10") %>% 
  filter(matrix != "Nosenko2013_ribo_11057") %>% 
  filter(matrix != "Borowiec2015_Best108")

taxon_rectangles$matrix = 
  factor( 
    taxon_rectangles$matrix, 
    levels = c("Dunn2008", "Hejnol2009", "Philippe2009", "Nosenko2013_ribo_14615", "Ryan2013_est", 
             "Moroz2014_3d", "Whelan2015_D20", "Borowiec2015_Total1080", "Chang2015", 
             "Whelan2017_full", "Simion2017"))

taxon_rectangles %>%
  ggplot() +
    scale_x_continuous(name = "Gene sampling") +
    scale_y_continuous(name = "Taxa sampling") +
    geom_rect( mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = clade )) +
    scale_fill_manual(values = clade_colors, guide = guide_legend(reverse = TRUE)) +
    facet_wrap( ~ matrix ) + theme_classic()

@@

analyses_published_processed = 
  analyses_published %>%
  dplyr::filter( !grepl( "Recoding", model_combined ) ) %>%
  dplyr::rename( model = model_combined ) %>%
  select( model, clade, final, inference, manuscript, matrix ) %>%
  dplyr::rename( result = final ) %>%
  mutate( status = "published" )

@@

analyses_processed = 
  rbind( analyses_published_processed, analyses_new_processed ) %>%
  # mutate(software=recode( software, `PhyloBayes MPI`="PhyloBayes") ) %>%
  mutate(model = recode(model,
                    `GTR20` = "GTR",
                    `CAT+F81` = "Poisson+CAT",
                    `Poisson + CAT` = "Poisson+CAT",
                    `GTR + CAT` = "GTR+CAT")) %>%
  mutate( 
    model = factor( model, 
      levels = c("WAG", "LG", "GTR", "data partitioning", 
                 "Poisson+C60", "WAG+C60", "LG+C60", "Poisson+CAT", "GTR+CAT") )) %>%
  mutate( status = factor(status, levels = c("published", "new")))

p = analyses_processed %>%
  ggplot( aes(x = model, y = clade, col = result, shape = inference) ) +
    facet_wrap( ~ status ) +
    geom_jitter( width = 0.20, height = 0.20, alpha = 0.7, size = 2 ) +
    scale_colour_manual(values = result_colors) +
    theme_classic() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    xlab("Model") + 
    ylab("Clade")

@@

# Simion2017 is the only Holozoa analysis to support Porifera-sister
analyses_processed %>% 
  filter(status == "published") %>% 
  filter(clade == "Holozoa") %>% 
  filter(result == "Porifera-sister")

# Check out the CAT analyses that supported Ctenophora sister
analyses_processed %>% 
  filter(status == "published") %>% 
  filter(result == "Ctenophora-sister") %>% 
  filter( grepl("CAT", model))

analyses_processed %>% 
  filter(status == "published") %>% 
  filter(result == "Ctenophora-sister") %>% 
  filter( grepl("CAT", model)) %>% 
  filter( clade != "Opisthokonta")

@@

# See results (and other fields) for new Cxx analyses, all are Ctenophora-sister with exception of Moroz2014_3d and Nosenko matrices
analyses_processed %>% 
  filter(status == "new") %>% 
  filter( grepl("C\\d\\d", model) ) %>% 
  print(n = 40)

@@

Philippe2009 = 
  analyses_sensitive %>% 
  filter(matrix == "Philippe2009_only_choanozoa" )

Philippe2009$model = 
  factor(
    Philippe2009$model, 
    levels = c("Poisson+CAT60", "Poisson+CAT90", "Poisson+CAT120", "Poisson+CAT150", "Poisson+CAT180"))

p1 = Philippe2009 %>% 
  ggplot( aes(model) ) +
  geom_line( aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group = 1 ), color = ("#E69F00")) + 
  geom_point( aes(y = `support_ctenophora_sister` ), color = ("#E69F00")) + 
  geom_line( aes(y = `support_porifera_sister`, colour = "Porifera-sister", group = 2), color = ("#009E73")) + 
  geom_point( aes(y = `support_porifera_sister` ), color = ("#009E73")) +
  scale_fill_manual(values = c("#E69F00", "#ebd196")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Number of CAT substitutional categories", y = "Support") + 
  ggtitle("Philippe2009_Choanimalia")

Philippe2009 = 
  analyses_sensitive %>% 
  filter(matrix == "Philippe2009_only_Holozoa")

Philippe2009$model = 
  factor(
    Philippe2009$model, 
    levels = c("Poisson+CAT270", "Poisson+CAT360", "Poisson+CAT400", "Poisson+CAT440", "Poisson+CAT480"))

p2 = Philippe2009 %>% 
  ggplot( aes(model)) +
  geom_line(aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group = 1), color = ("#E69F00")) + 
  geom_point(aes(y = `support_ctenophora_sister` ), color = ("#E69F00")) +
  geom_line(aes(y = `support_porifera_sister`, colour = "Porifera-sister", group = 2), color = ("#009E73")) + 
  geom_point(aes(y = `support_porifera_sister` ), color = ("#009E73")) +
  scale_fill_manual(values = c("#E69F00", "#ebd196")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Number of CAT substitutional categories", y = "Support") + 
  ggtitle("Philippe2009_Holozoa")

Whelan2017_full_only_choanozoa = 
  analyses_sensitive %>% 
  filter(matrix == "Whelan2017_full_only_choanozoa")

Whelan2017_full_only_choanozoa$model = 
  factor( 
    Whelan2017_full_only_choanozoa$model, 
    levels = c("Poisson+CAT60", "Poisson+CAT340", "Poisson+CAT380", "Poisson+CAT420", "Poisson+CAT460"))

p3 = Whelan2017_full_only_choanozoa %>% 
  ggplot( aes(model)) +
  geom_line(aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group = 1), color = ("#E69F00")) + 
  geom_point(aes(y = `support_ctenophora_sister` ), color = ("#E69F00")) +
  geom_line(aes(y = `support_porifera_sister`, colour = "Porifera-sister", group = 1), color = ("#009E73")) + 
  geom_point(aes(y = `support_porifera_sister` ), color = ("#009E73")) +
  scale_fill_manual(values = c("#E69F00", "#ebd196")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Number of CAT substitutional categories", y = "Support") + 
  ggtitle("Whelan2017_full_Choanozoa")

Whelan2017_strict = 
  analyses_sensitive %>% 
  filter(matrix == "Whelan2017_strict" )

Whelan2017_strict$model = 
  factor( 
    Whelan2017_strict$model, 
    levels = c("Poisson+CAT60", "Poisson+CAT70", "Poisson+CAT80", "Poisson+CAT90", "Poisson+CAT100",
             "Poisson+CAT110", "Poisson+CAT120", "Poisson+CAT150", "Poisson+CAT180", "Poisson+CAT360"))

p4 = Whelan2017_strict %>% 
  ggplot( aes(model)) +
  geom_line(aes(y = `support_ctenophora_sister`, colour = "Ctenophora-sister", group = 1), color = ("#E69F00")) + 
  geom_point(aes(y = `support_ctenophora_sister` ), color = ("#E69F00")) +
  geom_line(aes(y = `support_porifera_sister`, colour = "Porifera-sister", group = 1), color = ("#009E73")) + 
  geom_point(aes(y = `support_porifera_sister` ), color = ("#009E73")) +
  scale_fill_manual(values = c("#E69F00", "#ebd196")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Number of CAT substitutional categories", y = "Support") + 
  ggtitle("Whelan2017_strict")

ggarrange(ggarrange(p1, p2, ncol = 2, labels = c("A", "B")), nrow = 3,
          p3, p4, labels = c("C", "D"))
```
**Fig. 4.**  Sensitive analyses with representative data matrices analyzed by different number of substitutionary categories (nCAT) in PhyloBayes. Statistical support values (posterior probabilities) were obtained from three data matrcies using the site-heterogeneous Poisson-CAT model with different categories” (A) Phlippe2009_Choanozoa; (B) Philippe2009_Holozoa; (C) Whelan2017_full_Choanozoa; (D) Whelan2017_strict. Statistical support for Ctenophora-sister and Porifera-sister is indicated in orange and green, respectively. Support values are from the sensitive analyses are shown in Supplementary table XX.

@@

**Extended Data Table 2.** The models selected by ModelFinder for each matrix.
```{r table_modelfinder}
analyses_new %>%
  filter(modelfinder == TRUE) %>%
  select( matrix, clade, result, model_summary ) %>%
  kable()

@@

matrix_overlap_rectangles =
  lapply( 
    sequence_matrices[no_subsets], 
    function(x) lapply(sequence_matrices[no_subsets], function(y) overlap_rects(x, y))) %>%
  unlist(recursive = FALSE) %>%
  bind_rows()

matrix_overlap_rectangles %>% 
  ggplot( ) +
    scale_x_continuous(name = "Genes") +
    scale_y_continuous(name = "Species") +
    geom_rect( mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = MSA ), alpha = 0.5) +
    facet_grid( matrix_1 ~ matrix_2, scales = "free" ) +
    theme( axis.text.x = element_text(angle = 45, hjust = 1), 
           strip.text.y = element_text(angle = 0), 
           strip.text.x = element_text(angle = 60))

@@

ribo$matrix = 
  factor( 
    ribo$matrix, 
    levels = c("Dunn2008", "Philippe2009", "Hejnol2009", "Nosenko2013_nonribo_9187", "Nosenko2013_ribo_11057", 
             "Nosenko2013_ribo_14615", "Ryan2013_est", "Moroz2014_3d", "Chang2015", "Borowiec2015_Best108", 
             "Borowiec2015_Total1080", "Whelan2015_D1", "Whelan2015_D10", "Whelan2015_D20", "Simion2017", 
             "Whelan2017_full", "Whelan2017_strict") )

p1 = ribo %>% 
  ggplot( aes(x = matrix, y = number_of_genes, fill = gene_category) ) + 
  geom_bar(stat = "identity") + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Matrix", y = "Number of genes") + 
  scale_fill_manual(values = c("grey", "black"), labels = c("Non-ribosomal genes", "Ribosomal genes"))

busco$matrix = 
  factor( 
    busco$matrix, 
    levels = c("Dunn2008", "Philippe2009", "Hejnol2009", "Nosenko2013_nonribo_9187", "Nosenko2013_ribo_11057", 
               "Nosenko2013_ribo_14615", "Ryan2013_est", "Moroz2014_3d", "Chang2015", "Borowiec2015_Best108", 
               "Borowiec2015_Total1080", "Whelan2015_D1", "Whelan2015_D10", "Whelan2015_D20", "Simion2017", 
               "Whelan2017_full", "Whelan2017_strict") )

busco$gene_category = factor(busco$gene_category, levels = c("non_busco", "busco"))

p2 = busco %>% 
  ggplot( aes(x = matrix, y = number_of_genes, fill = gene_category) ) + 
  geom_bar(stat = "identity") + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Matrix", y = "Number of genes") + 
  scale_fill_manual( values = c("grey", "black"), labels = c("Non-BUSCO genes", "BUSCO genes") )

@@

```{r figure_phylogenetic_signal, fig.width=6, fig.height=12}

au_tests %>%
  unite(Matrix, c(manuscript, matrix, model)) %>%
  group_by(Matrix, result) %>%
  count() %>%
  group_by(Matrix) %>%
  mutate(percent = prop.table(n)) %>%
  ggplot(aes(x = Matrix, y = percent, fill = result)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("#E69F00", "#ebd196", "#a2e8d5", "#009E73")) + 
    coord_flip() +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank())

```
**Extended Data Fig. 3** The distribution of phylogenetic signal for two alternative topological hypotheses on the earliest-branching animal lineage with different models and outgroup choice in Philippe2009, Ryan2013 and Whelan2017_full matrices (with two outgroup sampling: Choanozoa and full matrices). The two alternative topological hypotheses are: Ctenophora-sister; T1 (Orange);Porifera-sister T2 (Green). By calculating gene-wise log-likelihood scores between T1 and T2 for every gene, we considered a gene with a log-likelihood difference of two as a gene with strong (lnl > 2) or weak (lnl < 2) phylogenetic signal for each hypothesis (Smith et al. 2020).
